Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-09-03 Manuscript/manuscript.Rmd
added refs in intro

A key factor in locating TAD boundaries is Hi-C data "resolution". Resolution determines the bin sizes of the linear genome used to enumerate contacts caused by chromatin folding. Lower resolutions contribute to larger bin sizes, which lead to variability in the location of TAD boundaries. Due to high sequencing costs, most available Hi-C datasets have relatively low resolution, ranging from 25 kb to 100 kb. Because increasing the resolution of Hi-C data requires a quadratic increase in the total sequencing depth, obtaining high resolution remains difficult [@schmitt2016genome]. While these low resolution Hi-C datasets can be used to probe higher order genomic patterns such as A/B compartments, more refined resolutions are necessary to accurately define sub-domains and the boundaries that demarcate them. Furthermore, even at its highest (750 bp [@bonev2017multiscale]), Hi-C sequencing resolution still remains well below the resolution of ChIP-Sequencing experiments used to annotate the genome, which ranges in the tens of base pairs. This resolution gap between 1D and 3D genomic maps can make it difficult to disentangle the role of genomic factors related to TAD boundaries, thereby limiting our understanding of epigenomic regulation and its potential link to chromatin architecture. We hypothesize that, by leveraging genomic annotations--defined by high resolution ChIP-Seq data--known to be enriched at TAD boundaries, we may be able more to precisely predict the locations of TAD boundaries at higher resolutions.

>>>>

A key factor in locating TAD boundaries is "resolution". Due to the high-dimensional nature of Hi-C data, genome binning has been a common strategy for dealing with a low signal-to-noise ratio problem [@li2019comparison]. Resolution, therefore, determines the bin sizes of the linear genome used to enumerate contacts caused by chromatin folding. Lower resolutions contribute to larger bin sizes, which lead to variability in the location of TAD boundaries. Due to high sequencing costs, most available Hi-C datasets have relatively low resolution, ranging from 25 kb to 100 kb. Because increasing the resolution of Hi-C data requires a quadratic increase in the total sequencing depth, obtaining high resolution remains difficult [@schmitt2016genome]. While these low resolution Hi-C datasets can be used to probe higher order genomic patterns such as A/B compartments, more refined resolutions are necessary to accurately define sub-domains and the boundaries that demarcate them [@fudenberg2016formation]. Furthermore, even at its highest (750 bp [@bonev2017multiscale]), Hi-C sequencing resolution still remains well below the resolution of ChIP-Sequencing experiments used to annotate the genome, which ranges in the tens of base pairs [@mendoza2009genome]. This resolution gap between 1D and 3D genomic maps can make it difficult to disentangle the role of genomic factors related to TAD boundaries, thereby limiting our understanding of epigenomic regulation and its potential link to chromatin architecture. We hypothesize that, by leveraging genomic annotations--defined by high resolution ChIP-Seq data--known to be enriched at TAD boundaries, we may be able more to precisely predict the locations of TAD boundaries at higher resolutions.
